### 2015-11-10 reply2referee.txt
responding to referee's comments. reply2referee.txt includes to-do list with three items

***** TO DO LIST *****

1) response to ``The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. ''

2) argue the necesity of ``Observability`` section, in particular S4.3 and S4.4.

3) response to ``it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider. ''

**********************

--------------------------------------------------------------------------------------
Comparison with previous work:

The nature of radio emission from red giant Hot Jupiters has already been addressed by Ignace et al (2010). The main difference in this new paper is that the authors explicitely allow for the area presented to the stellar wind by the exoplanetary magnetosphere, rather than simply the area presented by the bow shock around an unmagnetised planet. This provides around two orders of magnitude increase in the input kinetic power intercepted by the exoplanet.

This approach requires that the authors make some estimate of the magnetic field strength of the planet, for which they use published expressions for the field strength as a function of the exoplanet's mass and age. As a smaller point, in the present paper, the authors also explicitely consider the ionisation of the plasma around the exoplanet due to simple Bondi-Hoyle accretion. In the Ignace paper this mechanism was also introduced, but not explored in detail as the ionisation was provided by the presence of the bow shock that forms around the exoplanet.

The role of the size of the exoplanetary magnetosphere has been addressed by many other authors, including many of the papers by Zarka and Griessmeier cited by the authors. The novel aspect is the application to red giants. Most of the other parts of the paper involve applying published results. The most interesting of these is the use of the published expressions for the drop in the planet's convective heat flux with time as a means of estimating the decay with time of the planetary field strength and hence the emission frequency.
--------------------------------------------------------------------------------------

We thank referee's pointing this out, which made us realize that we previously failed to put our work into the context. 
As referee understands, the main topic of our paper is the re-evaluation of radio emission from planets around evolved stars. While the possibility of their strong radio emission has been addressed in Ignace et al. (2011), there are important updates in this paper, in particular:

- taking account of the shrinking of planetary magnetosphere. As a result, our estimate of radio emission is order-of-magnitude off of their estimate.  
- pointing out the presence of ionizing radition from accretion and quantitative evaluation of their ionizing effects. A mechanism to ionize otherwise neutral stellar wind has been briefly introduced in Ignace et al. (2011), but it is different from the mechanism we introduce. 
- taking account of the plasma frequency of stellar wind, which severely affects the detectability of radio signal.

In the revised manuscript, we tried to make these point clearer in Introduction section. The revised parts are marked in bold font.

Previously, we covered some topics that have already been addressed by previous papers, with an intention of making the paper self-contained and easy to follow. According to referee's suggestion, we attempted to reduce the materials as much as we think reasonable <<<grammatically wrong?>>>.  
There are some arguments which we still think are worth part of the paper. We describe them in the following, in reponse to each of referee's detailed comments.

--------------------------------------------------------------------------------------
Detailed comments:

Introduction: The authors should cite Jardine & Cameron (2008) who provide a model for the radio emission that does not rely on scaling of the solar system, and Smith et al 2009 for a search for a secondary eclipse.
--------------------------------------------------------------------------------------

We now included there papers in this section.

--------------------------------------------------------------------------------------
Section 2.2

Eqns 6-8 seem to be missing a factor of 4pi.
--------------------------------------------------------------------------------------

We thank the referee for pointing this out. We think we missed pi for equations 7-8 instead of 4pi, since we are interested in the cross-sectional area of the stellar wind energy input. We believe eqn 6 does not need 4pi.

--------------------------------------------------------------------------------------
Section 2.3

This section is only needed to calculate the decay with time of the field strength of the planet. For the radio emission, the authors simply scale to a field strength of 10G (which is perfectly acceptable as most of the parameters used in section 2.4 are unknown). This section simply quotes results from other papers and could be moved to an appendix.
--------------------------------------------------------------------------------------

The model of planetary magnetic field descibed in Section 2.3 is necessary to discuss the planetary parameter space, especially in terms of planetary mass and orbit, shown in Figure 3. Therefore, we would like to retain the description in the main manuscript. 
On the other hand, it is true that the scaling of planetary magnetic field is not a novel point of this paper (as explicitly mentioned in the manuscript), so we removed some panels from figure 2.

--------------------------------------------------------------------------------------
Section 2.4

The same comment applies here...the authors ultimately only use a wind velocity that is 1/10 that of the solar wind, so much of this section is unnecessary.
--------------------------------------------------------------------------------------

It is true that some of the points discussed here are not new points. Nevertheless, we think it is our responsibility to clarify our assumptions and that readers would be rather benefited from these descriptions because these properties are not common sense in the field of exoplanets. 
It is necessary to quantitatively evaluate not only the velocity but also significant mass-loss rate. Also, the typical values of stellar wind density is helpful and necessary step to discuss plasma frequency.

--------------------------------------------------------------------------------------
Section 2.5

This section is new, but not necessary, as Ignace (2010) have already addressed the issue of the ionisation of the exoplanetary exosphere. 
--------------------------------------------------------------------------------------

In Ignace et al. (2010) paper, post-shock ionization was introduced. The mechanism we newly propose in this paper provides a different view, and this allows us to adopt the same scaling law of radio emission as that of Solar-System planets. Therefore we think this is an important part of our work and we will not remove this from the paper.

--------------------------------------------------------------------------------------
Section 3.2

Given that the authors dismiss the AGB stars because their cutoff frequencies are too high (which could have been demonstrated very briefly) it seems unnecessary to include these at every step. More significantly, the comparison with the "classical" hot Jupiters is not valid, either in eqns 43 or 48, as in this case the orbital velocity of the planet cannot be neglected when considering the relative velocity of the stellar wind particles to the planet (equation 7). The stellar rotational velocity is probably small and so can be neglected however. The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. 
--------------------------------------------------------------------------------------

As for the AGB stars, we believe it is important to discuss not only potentially detectable signals but also the limitation of our ability of detection. In addition, there are some parameter range where AGB companions could be detected. Therefore, we think it is worth including the discussion of AGB stars.

As for the comparison with "classical" hot Jupiters, we appreciate referee's comments. We estimated the orbital velocity of our ``canonical hot Jupiter'' at 0.05 AU to be ~130 km/sec, which can contribute 5% increase in the relative velocity. We mentioned in the paragraph following equation (43).

<<< ram pressure ??? >>>

--------------------------------------------------------------------------------------
Fig 3: the numbers on the RHS have no units...they are probably fluxes, but they should be shown. 
--------------------------------------------------------------------------------------

We appreciate this feedback. We have indicated the unit in the title of Figure 3.

--------------------------------------------------------------------------------------
Section 4:

Much of this section could be cut, saving perhaps section 4.2. Predictions for observations with different facilities have been made many times before.
--------------------------------------------------------------------------------------

We would like to retain Section 4.1; indeed, when we talked about this work with people, they often ask me about it. 
We also think Section 4.3 and 4.4 are useful. <<< To Tony, please describe how difficult (or not straight-forward) it is to obtain these number >>>

--------------------------------------------------------------------------------------
Section 5 and Appendix B:

Could be cut.
--------------------------------------------------------------------------------------

Following the referee's comment, we cut Section 5.1, while some part of it are merged with the last section (shown in bold font). 
The intention of Section 5 was to answer a potential concern about the canceling field of far more massive stellar wind than sollar wind. We have clarified this motivation and moved it to Appendix. 
Previously we thought Appendix B might be useful for future workers (since we often got confused when reading previous papers), but we decided to cut it for the sake of brevity, as the referee suggested.

--------------------------------------------------------------------------------------
Section 6:

The authors note the rareness of evolved stars - it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider.
--------------------------------------------------------------------------------------

As far as we understand the referee's feedback, Figure 5 partly answers this suggestion, where the expected flux from the known system is estimated. We now improved the discussion by including the effects of occurrence rate of Jupiter-like planets. <<< to do >>>