### 2015-11-11 reply2referee.txt
Reformat response file; some minor wording changes; add cover letter to editor.

***** TO DO LIST *****

1) response to ``The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. ''

2) argue the necesity of ``Observability`` section, in particular S4.3 and S4.4.

3) response to ``it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider. ''

**********************

                                                10 November 2015

Dear Dr. Burton,

  We now resubmit our paper titled "Radio Emission from Red Giant Hot
Jupiters" by Yuka Fujii and collaborators, having addressed the
referee's points.  In so doing we have provided additional explanatory
text and updated the references, both in ways suggested by the referee
and on our own initiative.  Below is our point-by-point response to
the referee.  In our response, we present the referee's comments
preceded by "## ", and our own comments indented in the space that
follows.

  We realized upon reading the referee's report that we should have
done a better job of motivating some of the sections in our paper.
For instance, the section about the canceling field potentially
generated by stellar-wind electrons spiraling down field lines as they
accrete onto the planet is precisely in order to allay a reader's
possible concerns that the planetary magnetic field would be
significantly attenuated by the canceling field generated by the
massive stellar wind from an AGB star, a possibility that we wrestled
with for probably 100 hours.  We hope to save our readers from having
to repeat the same thought process we went through by laying out the
logic clearly.  However, it seems that we didn't explain well enough
why we put this section in the paper; in this resubmission, we have
tried to better motivate the parts that the referee thought were less
germane.

  In addition, it appears that the referee and we have somewhat
different philosophies regarding the purpose of a paper.  For example,
the referee suggests that we eliminate the section "Assumptions for
planetary magnetic field" and instead simply state that we assume a
magnetic field of 10G.  In our view, this would significantly weaken
our manuscipt.  Our aim is not merely to explain the assumptions that
go into making each figure and table but also to walk the reader
through the physics of the processes under discussion.

  Finally, we point out that this comment from the report was
surprising:
----
  "Section 2.5 This section is new, but not necessary, as Ignace
   (2010) have already addressed the issue of the ionisation of the
   exoplanetary exosphere."
----
Our manuscript presents a more detailed and accurate description of
the ionisation profile around a planet accreting stellar wind than
that in Ignace et al.'s paper.  It is unusual, in our experience, for
a referee to say that a description in a paper that is new, correct,
and more detailed than another description in the literature, is
unnecessary.

  Should you have any further questions, please feel free to get in
touch with the corresponding author, Yuka Fujii.

All the very best,
  Yuka Fujii (for the authors)

--------------------------------------------------------------------------------------

## Reviewer's Comments:
##
## This paper addresses the possibility of observing radio emission from
## planets orbiting around red giants. This emission is powered by the
## interaction of the stellar wind with the exoplanet, in a manner
## similar to that observed from many solar-system planets including
## Jupiter and the Earth. The winds of red giants are much cooler and
## slower than the solar wind, but they have a much higher mass
## density. Their kinetic energy fluxes therefore may be large enough to
## power observable radio fluxes. There are currently many planned and
## ongoing searches for radio emission from exoplanets, so this is a
## timely topic to address.
## 
##
## Comparison with previous work: 
##
## The nature of radio emission from red giant Hot Jupiters has already
## been addressed by Ignace et al (2010). The main difference in this new
## paper is that the authors explicitely allow for the area presented to
## the stellar wind by the exoplanetary magnetosphere, rather than simply
## the area presented by the bow shock around an unmagnetised
## planet. This provides around two orders of magnitude increase in the
## input kinetic power intercepted by the exoplanet.
##
## This approach requires that the authors make some estimate of the
## magnetic field strength of the planet, for which they use published
## expressions for the field strength as a function of the exoplanet's
## mass and age. As a smaller point, in the present paper, the authors
## also explicitely consider the ionisation of the plasma around the
## exoplanet due to simple Bondi-Hoyle accretion. In the Ignace paper
## this mechanism was also introduced, but not explored in detail as the
## ionisation was provided by the presence of the bow shock that forms
## around the exoplanet.
##
## The role of the size of the exoplanetary magnetosphere has been
## addressed by many other authors, including many of the papers by Zarka
## and Griessmeier cited by the authors. The novel aspect is the
## application to red giants. Most of the other parts of the paper
## involve applying published results. The most interesting of these is
## the use of the published expressions for the drop in the planet's
## convective heat flux with time as a means of estimating the decay with
## time of the planetary field strength and hence the emission frequency.

  We thank the referee for pointing out that we were not as clear as
  we would like to have been regarding the many differences from and
  advances upon the seminal paper of Ignace (2010) in our original
  submission.  In this resubmission, we have taken the referee's
  comments to heart and have endeavored to trim literature review
  where it is unnecessary, and to highlight the new aspects of our
  work.

## Recommendation:
## 
## My main concern with this paper is that it is much too long for its
## original scientific content. The amendment to the Ignace (2010) result
## is interesting, as is the decay with time of the emission frequency,
## but much of the rest is unnecessary. For example, Appendix B covers
## almost two pages, but is not referred to or used in the paper. I have
## outlined below detailed comments on each section, but my
## recommendation is that the authors reduce the main paper by a factor
## of four by focussing on only the new and relevant results.

  We thank referee for this comment.  We now realize that we had
  previously failed to put our work into proper context.  [DSS: strike
  the following: As referee understands,] the main topic of our paper
  is the re-evaluation of radio emission from planets around evolved
  stars. While the possibility of their strong radio emission has been
  addressed in Ignace et al. (2011) [DSS: don't we mean 2010?], there
  are important updates in this paper, in particular:

  - taking account of the shrinking of planetary magnetosphere. As a
    result, our estimate of radio emission is order-of-magnitude off
    of their estimate.
  - pointing out the presence of ionizing radition from accretion and
    quantitative evaluation of their ionizing effects. A mechanism to
    ionize otherwise neutral stellar wind has been briefly introduced
    in Ignace et al. (2011), but it is different from the mechanism we
    introduce.
  - taking account of the plasma frequency of stellar wind, which
    severely affects the detectability of radio signal.

  In the revised manuscript, we tried to make these point clearer in
  Introduction section. The revised parts are marked in bold font.

  Previously, we covered some topics that have already been addressed
  by previous papers, with an intention of making the paper
  self-contained and easy to follow. According to referee's
  suggestion, we attempted to reduce the materials as much as we think
  reasonable <<<grammatically wrong?>>>.  There are some arguments
  which we still think are worth part of the paper. We describe them
  in the following, in reponse to each of referee's detailed comments.

## Detailed comments:
## 
## Introduction: The authors should cite Jardine & Cameron (2008) who
## provide a model for the radio emission that does not rely on scaling
## of the solar system, and Smith et al 2009 for a search for a secondary
## eclipse.

  We now included their papers in this section.

## Section 2.2
## 
## Eqns 6-8 seem to be missing a factor of 4pi.

  This is a good catch by the referee!  It looks like we missed a
  factor of pi in equations 7-8 instead of 4pi, since we are
  interested in the cross-sectional area of the stellar wind energy
  input.  Equation 6 looks like it's okay since it's a
  proportionality, not equality.

## Section 2.3
## 
## This section is only needed to calculate the decay with time of the
## field strength of the planet. For the radio emission, the authors
## simply scale to a field strength of 10G (which is perfectly acceptable
## as most of the parameters used in section 2.4 are unknown). This
## section simply quotes results from other papers and could be moved to
## an appendix.

  The model of planetary magnetic field descibed in Section 2.3 is
  necessary to discuss the planetary parameter space, especially in
  terms of planetary mass and orbit, shown in Figure 3. Therefore, we
  would like to retain the description in the main manuscript.  On the
  other hand, it is true that the scaling of planetary magnetic field
  is not a novel point of this paper (as explicitly mentioned in the
  manuscript), so we removed some panels from figure 2.

## Section 2.4
## 
## The same comment applies here...the authors ultimately only use a wind
## velocity that is 1/10 that of the solar wind, so much of this section
## is unnecessary.

  It is true that some of the points discussed here are not new
  points. Nevertheless, we think it is our responsibility to clarify
  our assumptions and that readers would be rather benefited from
  these descriptions because these properties are not common sense in
  the field of exoplanets.  It is necessary to quantitatively evaluate
  not only the velocity but also significant mass-loss rate. Also, the
  typical values of stellar wind density is helpful and necessary step
  to discuss plasma frequency.

## Section 2.5
## 
## This section is new, but not necessary, as Ignace (2010) have
## already addressed the issue of the ionisation of the exoplanetary
## exosphere.

  In the Ignace et al. (2010) paper, post-shock ionization was
  introduced. The mechanism we newly propose in this paper provides a
  different view, and this allows us to adopt the same scaling law of
  radio emission as that of Solar-System planets. Therefore we think
  this is an important part of our work and we will not remove this
  from the paper.

## Section 3.2
## 
## Given that the authors dismiss the AGB stars because their cutoff
## frequencies are too high (which could have been demonstrated very
## briefly) it seems unnecessary to include these at every step. More
## significantly, the comparison with the "classical" hot Jupiters is not
## valid, either in eqns 43 or 48, as in this case the orbital velocity
## of the planet cannot be neglected when considering the relative
## velocity of the stellar wind particles to the planet (equation 7). The
## stellar rotational velocity is probably small and so can be neglected
## however. The authors should demonstrate that the ram pressure of any
## exoplanetary outflow induced by heating can also be neglected in
## determining the size of the magnetosphere.

  As for the AGB stars, we believe it is important to discuss not only
  potentially detectable signals but also the limitation of our
  ability of detection. In addition, there are some parameter range
  where AGB companions could be detected. Therefore, we think it is
  worth including the discussion of AGB stars.

  As for the comparison with "classical" hot Jupiters, we appreciate
  referee's comments. We estimated the orbital velocity of our
  ``canonical hot Jupiter'' at 0.05 AU to be ~130 km/sec, which can
  contribute 5% increase in the relative velocity. We mentioned in the
  paragraph following equation (43).

  <<< ram pressure ??? >>>

## Fig 3: the numbers on the RHS have no units...they are probably
## fluxes, but they should be shown.

  We appreciate this feedback. We have indicated the unit in the title
  of Figure 3.

## Section 4:
## 
## Much of this section could be cut, saving perhaps section
## 4.2. Predictions for observations with different facilities have been
## made many times before.

  We would like to retain Section 4.1; indeed, when we talked about
  this work with people, they often ask me about it.  We are unaware
  of any predictions in the literature that make it easy for the
  reader to estimate the required duration of the observations
  describe in our manuscript.  We also think Section 4.3 and 4.4 are
  useful. <<< To Tony, please describe how difficult (or not
  straight-forward) it is to obtain these number >>>

## Section 5 and Appendix B:
## 
## Could be cut.

  Following the referee's comment, we cut Section 5.1, while some part
  of it are merged with the last section (shown in bold font).  The
  intention of Section 5 was to answer a potential concern about the
  canceling field of far more massive stellar wind than sollar
  wind. We have clarified this motivation and moved it to Appendix.
  Previously we thought Appendix B might be useful for future workers
  (since we often got confused when reading previous papers), but we
  decided to cut it for the sake of brevity, as the referee suggested.

  [DSS: I don't think it should be moved to the Appendix.]

## Section 6:
## 
## The authors note the rareness of evolved stars - it would be useful if
## they included a brief calculation or statement of how many might be
## expected, particularly in the age ranges they consider.

  As far as we understand the referee's feedback, Figure 5 partly
  answers this suggestion, where the expected flux from the known
  system is estimated. We now improved the discussion by including the
  effects of occurrence rate of Jupiter-like planets. <<< to do >>>