### 2015-12-20 reply2referee.txt
Edits to reply2referee.txt for flow.

                                                12 December 2015

Dear Dr. Burton,

  We now resubmit our paper titled "Radio Emission from Red Giant Hot
Jupiters" by Yuka Fujii and collaborators, having addressed the
referee's points.  In so doing we have provided additional explanatory
text and updated the references, both in ways suggested by the referee
and on our own initiative.  Below is our point-by-point response to
the referee.  In our response, we present the referee's comments
preceded by "## ", and our own comments indented in the space that
follows.  We have also revised the reference list to match the
journal's nomenclature.

  We realized upon reading the referee's report that we should have
done a better job of motivating some of the sections in our paper.
For instance, we wrote the section about the "canceling field"
generated by stellar-wind electrons, as they spiral down field lines
in accreting onto the planet, in order to allay the reader's possible
concerns that the planetary magnetic field would be significantly
attenuated by the canceling field that would be generated by the
massive stellar wind from an AGB star -- a possibility that we
wrestled with for probably 100 hours.  We hope to save our readers
from having to repeat the same thought process we went through by
laying out the logic clearly.  However, it seems that we didn't
explain well enough why we put this section in the paper; in this
resubmission, we have tried to better motivate the parts that the
referee thought were less germane.

  In addition, it appears that the referee and we have somewhat
different philosophies regarding the purpose of a paper.  For example,
the referee suggests that we eliminate the section "Assumptions for
planetary magnetic field" and instead simply state that we assume a
magnetic field of 10G.  In our view, this would significantly weaken
our manuscipt.  Our aim is not merely to explain the assumptions that
go into making each figure and table but also to walk the reader
through the physics of the processes under discussion.

  Finally, we point out that this comment from the report was
surprising:
----
  "Section 2.5 This section is new, but not necessary, as Ignace
   (2010) have already addressed the issue of the ionisation of the
   exoplanetary exosphere."
----
Our manuscript presents a more detailed and accurate description of
the ionization profile around a planet accreting stellar wind than
that in Ignace et al.'s paper.  It is unusual, in our experience, for
a referee to say that a description that is new, correct, and more
detailed than another description in the literature, is unnecessary.

  Should you have any further questions, please feel free to get in
touch with the corresponding author, Yuka Fujii.

All the very best,
  Yuka Fujii (for the authors)

--------------------------------------------------------------------------------------

## Reviewer's Comments:
##
## This paper addresses the possibility of observing radio emission from
## planets orbiting around red giants. This emission is powered by the
## interaction of the stellar wind with the exoplanet, in a manner
## similar to that observed from many solar-system planets including
## Jupiter and the Earth. The winds of red giants are much cooler and
## slower than the solar wind, but they have a much higher mass
## density. Their kinetic energy fluxes therefore may be large enough to
## power observable radio fluxes. There are currently many planned and
## ongoing searches for radio emission from exoplanets, so this is a
## timely topic to address.
## 
##
## Comparison with previous work: 
##
## The nature of radio emission from red giant Hot Jupiters has already
## been addressed by Ignace et al (2010). The main difference in this new
## paper is that the authors explicitely allow for the area presented to
## the stellar wind by the exoplanetary magnetosphere, rather than simply
## the area presented by the bow shock around an unmagnetised
## planet. This provides around two orders of magnitude increase in the
## input kinetic power intercepted by the exoplanet.
##
## This approach requires that the authors make some estimate of the
## magnetic field strength of the planet, for which they use published
## expressions for the field strength as a function of the exoplanet's
## mass and age. As a smaller point, in the present paper, the authors
## also explicitely consider the ionisation of the plasma around the
## exoplanet due to simple Bondi-Hoyle accretion. In the Ignace paper
## this mechanism was also introduced, but not explored in detail as the
## ionisation was provided by the presence of the bow shock that forms
## around the exoplanet.
##
## The role of the size of the exoplanetary magnetosphere has been
## addressed by many other authors, including many of the papers by Zarka
## and Griessmeier cited by the authors. The novel aspect is the
## application to red giants. Most of the other parts of the paper
## involve applying published results. The most interesting of these is
## the use of the published expressions for the drop in the planet's
## convective heat flux with time as a means of estimating the decay with
## time of the planetary field strength and hence the emission frequency.

  We thank the referee for pointing out that we were not as clear as
  we would like to have been regarding the many differences from and
  advances upon the seminal paper of Ignace (2010) in our original
  submission.  In this resubmission, we have taken the referee's
  comments to heart and have endeavored to trim literature review
  where it is unnecessary, and to highlight the new aspects of our
  work.

## Recommendation:
## 
## My main concern with this paper is that it is much too long for its
## original scientific content. The amendment to the Ignace (2010) result
## is interesting, as is the decay with time of the emission frequency,
## but much of the rest is unnecessary. For example, Appendix B covers
## almost two pages, but is not referred to or used in the paper. I have
## outlined below detailed comments on each section, but my
## recommendation is that the authors reduce the main paper by a factor
## of four by focussing on only the new and relevant results.

  We thank referee for this comment.  We now realize that we had
  previously failed to put our work into proper context.  The main topic
  of our paper is the re-evaluation of radio emission from planets around
  evolved stars. While the possibility of their strong radio emission has
  been addressed in Ignace et al. (2010), there are important updates in 
  this paper, in particular:

  - pointing out the presence of ionizing radiation from accretion and
    quantitatively evaluating the ionizing effects. A mechanism to
    ionize otherwise neutral stellar wind has been briefly introduced
    in Ignace et al. (2010), but it is different from the mechanism we
    introduce.

  - taking account of the shrinking of the planetary magnetosphere. As
    a result, our estimate of radio emission is an order-of-magnitude
    off of their scaling (eq. 20 of their paper).

  - taking account of the plasma frequency of stellar wind, which
    severely affects the detectability of radio signal.

  In the revised manuscript, we tried to make these points clearer in
  the Introduction section.  The revised parts are marked in bold and
  blue font.

  Previously, we covered some topics that have already been addressed
  by previous papers, with an intention of making the paper
  self-contained and easy to follow.  Based on the referee's
  suggestions, we attempted to reduce the amount of background
  material as much as we think reasonable.  There are some arguments
  from the literature that we still think are worth being part of the
  paper.  We describe them below, in reponse to the relevant parts of
  the referee's detailed comments.

## Detailed comments:
## 
## Introduction: The authors should cite Jardine & Cameron (2008) who
## provide a model for the radio emission that does not rely on scaling
## of the solar system, and Smith et al 2009 for a search for a secondary
## eclipse.

  Good suggestion.  We now include their papers in this section.

## Section 2.2
## 
## Eqns 6-8 seem to be missing a factor of 4pi.

  This is a good catch by the referee!  It looks like we missed a
  factor of pi in equations 7-8 instead of 4pi, since we are
  interested in the cross-sectional area of the stellar wind energy
  input.  Equation 6 looks like it's okay since it's a
  proportionality, not equality.

## Section 2.3
## 
## This section is only needed to calculate the decay with time of the
## field strength of the planet. For the radio emission, the authors
## simply scale to a field strength of 10G (which is perfectly acceptable
## as most of the parameters used in section 2.4 are unknown). This
## section simply quotes results from other papers and could be moved to
## an appendix.

  The model of planetary magnetic field descibed in Section 2.3 is
  necessary to discuss the planetary parameter space, especially in
  terms of planetary mass and orbit, shown in Figure 3.  Therefore, we
  would like to retain the description in the main manuscript.

## Section 2.4
## 
## The same comment applies here...the authors ultimately only use a wind
## velocity that is 1/10 that of the solar wind, so much of this section
## is unnecessary.

  It is true that some of the text in this section is discussion of
  the physics, and not new points/results.  Nevertheless, we think it
  is our responsibility to clarify our assumptions and that readers
  would benefit from these descriptions because these properties are
  not common sense in the field of exoplanets.  It is necessary to
  quantitatively evaluate not only the velocity but also the
  significant mass-loss rate.  Also, knowing the typical values of
  stellar wind density is helpful, and indeed a necessary step to
  discuss plasma frequency.

## Section 2.5
## 
## This section is new, but not necessary, as Ignace (2010) have
## already addressed the issue of the ionisation of the exoplanetary
## exosphere.

  In the Ignace et al. (2010) paper, post-shock ionization was
  introduced.  The mechanism we newly propose in this paper provides a
  different view, and this allows us to adopt the same scaling law of
  radio emission as that of Solar-System planets.  Therefore we think
  this is an important part of our work, and we believe that our
  manuscript is not improved by removing it.

## Section 3.2
## 
## Given that the authors dismiss the AGB stars because their cutoff
## frequencies are too high (which could have been demonstrated very
## briefly) it seems unnecessary to include these at every step. More
## significantly, the comparison with the "classical" hot Jupiters is not
## valid, either in eqns 43 or 48, as in this case the orbital velocity
## of the planet cannot be neglected when considering the relative
## velocity of the stellar wind particles to the planet (equation 7). The
## stellar rotational velocity is probably small and so can be neglected
## however. The authors should demonstrate that the ram pressure of any
## exoplanetary outflow induced by heating can also be neglected in
## determining the size of the magnetosphere.

  Regarding whether to include a discussion of AGB stars, we believe
  it is important to discuss not only potentially detectable signals
  but also the limitations of our ability to detect radio-emitting
  planets.  In addition, there are some parameter ranges where AGB
  companions could be detected.  Therefore, we think it is worth
  including the discussion of AGB stars.

  Regarding the comparison with "classical" hot Jupiters, we
  appreciate referee's comments.  The orbital velocity of our
  ``canonical hot Jupiter'' at 0.05 AU to be ~130 km/sec, which can
  contribute 5% increase in the relative velocity.  We now add the
  following text in the paragraph following equation (43):

    "We also note that the orbital velocity of the canonical hot
    Jupiters has been ignored in equation (45); the orbital velocity
    at 0.05~AU around a Solar-mass star is $\sim$130~km~sec$^{-1}$,
    which leads to only a $\sim$5\% increase of the relative velocity
    between planet and wind particles."

  We now add discussion about the effect of planetary outflow in
  Appendix A.

## Fig 3: the numbers on the RHS have no units...they are probably
## fluxes, but they should be shown.

  We appreciate this feedback.  We have indicated the units in the
  title of Figure 3.

## Section 4:
## 
## Much of this section could be cut, saving perhaps section
## 4.2. Predictions for observations with different facilities have been
## made many times before.

  We believe that Section 4.1 (regarding intrinsic radio emission from
  red giant stars) enhances the paper.  For instance, when we discuss
  this work with people, they often ask us about this topic.  It's a
  natural consideration, since a key component of observability is
  SNR, and the star obviously contributes some of the Noise (indeed,
  in most exoplanet contexts, the signal from the planet is competing
  against a much brighter star!).  To devote 1 page to what a reader
  might be likely to be concerned is the strongest source of Noise in
  the observations does not seem unreasonable to us.

  We also think Sections 4.3 and 4.4 are useful.  We are unaware of
  any predictions in the literature that make it easy for the reader
  to estimate the required duration of the observations describe in
  our manuscript.  Specifically, we know of no such comparison for
  state-of-the-art facilities that is up to date for late 2015, nor do
  we know of any consolidation of this information in one publication.
  Predictions for future facilities rapidly become out of date as
  projects are descoped (or "rebaselined" in the case of the SKA), and
  many sensitivity predictions for existing instruments were made
  before the instrument was built.  Including the latest measured
  performance or design specifications is crucial to assess observing
  feasibility.  In section 4.4, we discuss the important point that
  one can exploit the unique time variability of these signatures for
  detection.  We think this is a vital point to consider, as direct
  imaging would be impossible in the face of confusion from both
  Galactic and extragalactic sources (which are brighter but will not
  vary on the same time scale).

## Section 5 and Appendix B:
## 
## Could be cut.

  Following the referee's comment, we have cut Section 5.1, moving
  some parts of it to last section (shown in bold font) and other
  parts to the Appendix.  The intention of Section 5 was to answer a
  potential concern about the canceling field that would result from a
  stellar wind that is much denser than that of the Sun.  In preparing
  our original submission, we thought that the former Appendix B might
  be useful for future workers (since we often got confused when
  reading previous papers), but we decided to cut it for the sake of
  brevity, as the referee suggests.

## Section 6:
## 
## The authors note the rareness of evolved stars - it would be useful if
## they included a brief calculation or statement of how many might be
## expected, particularly in the age ranges they consider.

  Figure 5 partly answers this question, where the expected flux from
  the known system is estimated.  We have added a section to the
  Appendix (and quoted results from it in the main text) that
  explicitly addresses the referee's question in more detail by
  estimating how many potential targets are likely to be brighter than
  the detection limit.  In making this estimate, we take into account
  the fact that not all of stellar targets are likely to harbor the
  right kind of planet in the right orbital regime.